---
title: "NYC CitiBike System_EDA"
author: "S/17/811"
date: "2023-08-08"
output:
  word_document: default
  html_document:
    df_print: paged
---

***Exploratory Data Analysis***

```{r}
#import libraries

library(tidyverse)
library(lubridate)#'lubridate' package is used for working with dates and times in R
library(skimr)
library(tmap)#'tmap' package for creating thematic maps in R
library(sf)#'sf'(simple features) package is used for working with geospatial data in R

```

```{r}
#download data from "https://citibikenyc.com/system-data"

#import data

trip_data <- read_csv(file = "../data/trips_data_2022.csv",col_types =   c("cfTTcfcfddddf"))

head(trip_data)
```

```{r}
# explore the structure of the data set

trip_data %>% str()
```

```{r}
#summary statistic

trip_data %>% skim()
```

```{r}
#check missing values

#missmap(trip_data)
sum(is.na(trip_data))
```


```{r}
# filter observations which has the missing values

trip_data %>%
  filter(!complete.cases(.))
```

```{r}
# drop missing values

trip_data <- trip_data%>%drop_na()
```


```{r}
#Check duplicate rows

trip_data[duplicated(trip_data),]
```

```{r}
# split started_at and ended_at variables into two separate variables

trip_data <- trip_data %>%
  mutate(start_time = format(started_at, format = "%H:%M:%S"),
         start_date = as.Date(started_at),
         end_time = format(ended_at, format = "%H:%M:%S"),
         end_date = as.Date(ended_at)
  )
```


```{r}
#frequency distribution for categorical variables 'rideble_type', 'start_station_id', 'end_station_id', and 'member_casual'

frequency_ridebleType <- trip_data %>%
  group_by(rideable_type)%>%
  summarise(trips = n())%>%
  arrange(desc(trips))

frequency_ridebleType


frequency_startStation <- trip_data %>%
  group_by(start_station_id)%>%
  summarise(trips = n())%>%
  arrange(desc(trips))

frequency_startStation


frequency_endStation <- trip_data %>%
  group_by(end_station_id)%>%
  summarise(trips = n())%>%
  arrange(desc(trips))

frequency_endStation


frequency_member <- trip_data %>%
  group_by(member_casual)%>%
  summarise(trips = n())%>%
  arrange(desc(trips))

frequency_member
```

```{r}
# 'docked_bike' rideable type
trip_data %>% 
  filter(rideable_type == "docked_bike")
```

```{r}
frequency_docked_bike <- trip_data %>% 
  filter(rideable_type == "docked_bike")%>%
  group_by(start_station_id)%>%
  summarise(trips = n())%>%
  arrange(desc(trips))

frequency_docked_bike
```

```{r}
# remove observations where the rideable_type is 'docked_bike'

trip_data <- trip_data %>%
  filter(rideable_type != 'docked_bike')
```


```{r}
#Plot a bar chart of rideable types
trip_data %>%
  ggplot(aes(x = rideable_type)) +
  geom_bar()+
  labs(x = "Rideable Type", y="Trips",title = "Distribution of Rideable Type")+
  theme_classic()
```

```{r}
# distribution of trip counts for different 'rideable' types at each start station

trip_data %>%
  ggplot(mapping = aes(x=start_station_id))+
  geom_bar(aes(fill = rideable_type)) +
  xlab("start station") +
  ylab("trip count")
```

```{r}
# distribution of start stations

trip_data %>%
  ggplot(aes(x = start_station_id)) +
  geom_bar()+
  labs(x="Start Station ID",y="Trips",title = "Distribution of Start Station")+
  theme_classic()
```


```{r}
# distribution of end station
trip_data %>%
  ggplot(aes(x = end_station_id)) +
  geom_bar()+
  labs(x="End Station ID",y="Trips",title = "Distribution of End Station")+
  theme_classic()
```


```{r}
# User Behavior Patterns

# Identify top start and end stations for both member and casual users

top_start_stations_member <- trip_data %>%
  filter(member_casual == "member") %>%
  group_by(start_station_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(5)

top_start_stations_member

top_start_stations_casual <- trip_data %>%
  filter(member_casual == "casual") %>%
  group_by(start_station_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(5)

top_start_stations_casual

top_end_stations_member <- trip_data %>%
  filter(member_casual == "member") %>%
  group_by(end_station_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(5)

top_end_stations_member

top_end_stations_casual <- trip_data %>%
  filter(member_casual == "casual") %>%
  group_by(end_station_id) %>%
  summarise(count = n()) %>%
  arrange(desc(count)) %>%
  head(5)

top_end_stations_casual
```


Working with date and time distributions

```{r}
#distribution of number of trips by month 

# Extract month from start_date
#trip_data$month <- format(trip_data$start_date, "%B")

# Count the number of trips per month
#trip_counts <- trip_data %>%
#  group_by(month) %>%
#  summarise(total_trips = n())

# Create a bar plot of the number of trips by month
#ggplot(trip_counts, aes(x = month, y = total_trips)) +
#  geom_bar(stat = "identity") +
#  xlab("Month") +
#  ylab("Number of Trips") +
#  ggtitle("Distribution of Number of Trips by Month")+
#  theme_classic()
```


```{r}
# distribution of number of trips by day of the week

# Extract day of the week from start_date
trip_data$day_of_week <- weekdays(trip_data$start_date)

# Count the number of trips per day of the week
trip_counts <- trip_data %>%
  group_by(day_of_week) %>%
  summarise(total_trips = n())%>%
  arrange(desc(total_trips))  # Sort the data in descending order

# Reorder the levels of day_of_week based on total_trips
trip_counts$day_of_week <- factor(trip_counts$day_of_week, levels = trip_counts$day_of_week)

# Create a bar plot of the number of trips by day of the week
ggplot(trip_counts, aes(x = day_of_week, y = total_trips)) +
  geom_bar(stat = "identity") +
  xlab("Day of the Week") +
  ylab("Number of Trips") +
  ggtitle("Distribution of Number of Trips by Day of the Week")+
  theme_classic()
```


```{r}
# distribution of number of trips by day


# Extract day from start_date
#trip_data$day <- format(trip_data$start_date, "%d")

# Count the number of trips per day
#trip_counts <- trip_data %>%
#  group_by(day) %>%
#  summarise(total_trips = n())

# Create a bar plot of the number of trips by day
#ggplot(trip_counts, aes(x = day, y = total_trips)) +
#  geom_bar(stat = "identity") +
#  xlab("Day") +
#  ylab("Number of Trips") +
#  ggtitle("Distribution of Number of Trips by Day")+
#  theme_classic()
```


```{r}
# distribution of number of trips by hour

# Extract hour from start_time
trip_data$hour <- as.integer(substr(trip_data$start_time, 1, 2))

# Filter data for weekdays and weekends
weekday_trip_counts <- trip_data %>%
  filter(weekdays(start_date) %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) %>%
  group_by(hour) %>%
  summarise(total_trips = n())

weekend_trip_counts <- trip_data %>%
  filter(weekdays(start_date) %in% c("Saturday", "Sunday")) %>%
  group_by(hour) %>%
  summarise(total_trips = n())

# Create separate bar plots for weekdays and weekends
weekday_plot <- ggplot(weekday_trip_counts, aes(x = hour, y = total_trips)) +
  geom_bar(stat = "identity") +
  xlab("Hour") +
  ylab("Number of Trips") +
  ggtitle("Distribution of Number of Trips by Hour (Weekdays)") +
  theme_classic()

weekend_plot <- ggplot(weekend_trip_counts, aes(x = hour, y = total_trips)) +
  geom_bar(stat = "identity") +
  xlab("Hour") +
  ylab("Number of Trips") +
  ggtitle("Distribution of Number of Trips by Hour (Weekends)") +
  theme_classic()

# Display the plots
weekday_plot
weekend_plot

```

The weekdays plot provides insight into the distribution of the number of trips by hour on weekdays (Monday to Friday). Each bar on the x-axis represents a specific hour of the day, while the y-axis represents the count of trips recorded during each hour. By examining the heights of the bars, we can observe that the hours 8 AM, 3 PM, 4 PM, 5 PM, and 6 PM have the highest number of trips, indicating that these hours are the peak hours for weekday travel.

On the other hand, the weekends plot illustrates the distribution of the number of trips by hour on weekends (Saturday and Sunday). It follows a similar structure to the weekdays plot.By analyzing this plot, we can identify the hours 12 PM, 1 PM, 2 PM, 3 PM, 4 PM, and 5 PM as the peak hours for weekend travel.

Furthermore, we can observe that the count of trips during weekdays generally falls within the range of 0 to 700, whereas the count of trips during weekends tends to be lower, typically ranging from 0 to 225. This indicates that there are more trips recorded on weekdays compared to weekends.



```{r}
# Create box plots for the distribution of trips by hour, separated by weekdays and weekends

# Extract hour from start_time

trip_data$hour <- as.integer(substr(trip_data$start_time, 1, 2))

# Group the data by hour and count the number of trips

hourly_trip_counts <- trip_data %>%
  group_by(start_date,hour)%>%
  summarise(trip_count = n())

# Filter for weekdays(Monday to Friday)

weekday_trip_counts <- hourly_trip_counts%>% 
filter(weekdays(start_date) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))

# Filter for weekend days (Saturday and Sunday)

weekend_trip_counts <- hourly_trip_counts%>%
  filter(weekdays(start_date) %in% c("Saturday","Sunday"))

# Create box plots for weekdays
ggplot(weekday_trip_counts, aes(x = factor(hour), y = trip_count)) +
  geom_boxplot() +
  xlab("Hour of the Day") +
  ylab("Number of Trips") +
  ggtitle("Box Plot of Weekday Trip Counts by Hour") +
  theme_minimal() +
  scale_x_discrete(breaks = 0:23)

# Create box plots for weekend days
ggplot(weekend_trip_counts, aes(x = factor(hour), y = trip_count)) +
  geom_boxplot() +
  xlab("Hour of the Day") +
  ylab("Number of Trips") +
  ggtitle("Box Plot of Weekend Trip Counts by Hour") +
  theme_minimal() +
  scale_x_discrete(breaks = 0:23)

```


```{r}
# Create box plots for the distribution of trips by hour, separated by weekdays and weekends for user type(member_casual)

# Extract hour from start_time

trip_data$hour <- as.integer(substr(trip_data$start_time, 1, 2))

# Group the data by hour user type('member_casual') count the number of trips

hourly_trip_counts <- trip_data %>%
  group_by(start_date,hour,member_casual)%>%
  summarise(trip_count = n())

# Filter for weekdays(Monday to Friday)

weekday_trip_counts <- hourly_trip_counts%>% 
filter(weekdays(start_date) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))

# Filter for weekend days (Saturday and Sunday)

weekend_trip_counts <- hourly_trip_counts%>%
  filter(weekdays(start_date) %in% c("Saturday","Sunday"))

# Create box plots for weekdays, separated by user type

ggplot(weekday_trip_counts, aes(x = factor(hour), y = trip_count)) +
  geom_boxplot() +
  xlab("Hour of the Day") +
  ylab("Number of Trips") +
  ggtitle("Box Plot of Weekday Trip Counts by Hour") +
  theme_minimal() +
  scale_x_discrete(breaks = 0:23)+
  facet_wrap(~member_casual,scales = "free_y")

# Create box plots for weekend days, separated by user type

ggplot(weekend_trip_counts, aes(x = factor(hour), y = trip_count)) +
  geom_boxplot() +
  xlab("Hour of the Day") +
  ylab("Number of Trips") +
  ggtitle("Box Plot of Weekend Trip Counts by Hour") +
  theme_minimal() +
  scale_x_discrete(breaks = 0:23)+
  facet_wrap(~member_casual, scales = "free_y")
```


```{r}
# Create box plots for the distribution of trips by hour, separated by weekdays and weekends for bike type (classic_electric)

# Extract hour from start_time

trip_data$hour <- as.integer(substr(trip_data$start_time, 1, 2))

# Group the data by hour bike type('classic_electric') count the number of trips

hourly_trip_counts <- trip_data %>%
  group_by(start_date,hour,rideable_type)%>%
  summarise(trip_count = n())

# Filter for weekdays(Monday to Friday)

weekday_trip_counts <- hourly_trip_counts%>% 
filter(weekdays(start_date) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))

# Filter for weekend days (Saturday and Sunday)

weekend_trip_counts <- hourly_trip_counts%>%
  filter(weekdays(start_date) %in% c("Saturday","Sunday"))

# Create box plots for weekdays, separated by bike type

ggplot(weekday_trip_counts, aes(x = factor(hour), y = trip_count)) +
  geom_boxplot() +
  xlab("Hour of the Day") +
  ylab("Number of Trips") +
  ggtitle("Box Plot of Weekday Trip Counts by Hour") +
  theme_minimal() +
  scale_x_discrete(breaks = 0:23)+
  facet_wrap(~rideable_type,scales = "free_y")

# Create box plots for weekend days, separated by bike type

ggplot(weekend_trip_counts, aes(x = factor(hour), y = trip_count)) +
  geom_boxplot() +
  xlab("Hour of the Day") +
  ylab("Number of Trips") +
  ggtitle("Box Plot of Weekend Trip Counts by Hour") +
  theme_minimal() +
  scale_x_discrete(breaks = 0:23)+
  facet_wrap(~rideable_type, scales = "free_y")
```


```{r}
#Average, Maximum and Minimum Trip Duration

# Calculate trip duration and add as a new variable 'trip_duration' 

trip_data <- trip_data %>%
  mutate(trip_duration = difftime(ended_at, started_at, units = "mins"))

trip_data %>%
  summarise(maximum_trip_duration = max(trip_duration),
            minimum_trip_duration = min(trip_duration),
            average_trip_duration = mean(trip_duration))
```

```{r}
trip_data %>%
  filter(trip_duration == min(trip_duration) | trip_duration == max(trip_duration))
```


```{r}
#Convert trip_duration to numeric

trip_data$trip_duration <- as.numeric(as.character(trip_data$trip_duration))
```



```{r}
# Histogram of trip duration 
trip_data %>%
  ggplot(aes(x=trip_duration))+
  geom_histogram(binwidth = 5)+
  labs(title = "Histogram of Trip Durations",
       x = "Trip Duration (minutes)",
       y = "frequency")+
  theme_classic()
```


```{r}
# Create bins for trip duration and count the number of trips in each bin

trip_data %>%
  filter(trip_duration >= 0)%>%
  mutate(trip_duration_bin = cut(trip_duration,seq(0,max(trip_duration)+5,by =5),include.lowest = TRUE),right = FALSE)%>%
  group_by(trip_duration_bin)%>%
  summarise(num_trips = n())%>%
  arrange(trip_duration_bin)
```


```{r}
trip_data %>%
  filter(trip_duration >= 0)%>%
  mutate(trip_duration_bin = cut(trip_duration,seq(0,max(trip_duration)+5,by =5),include.lowest = TRUE),right = FALSE)%>%
  group_by(trip_duration_bin)%>%
  summarise(num_trips = n())%>%
  ggplot(aes(x = trip_duration_bin, y = num_trips))+
  geom_bar(stat = "identity")+
  labs(title = "Histogram of Trip Durations",
       x = "Trip Duration Bin",
       y = "Number of Trips")+
  theme_minimal()
```



```{r}
trip_data %>%
  ggplot(aes(x=" ",y=trip_duration))+
  geom_boxplot()+
  labs(title = "Box Plot of Trip Duration",y="Trip Duration (minutes)")+
  theme_minimal()
```


```{r}
# trip duration based on different users

user_type_summary <- trip_data %>%
  group_by(member_casual) %>%
  summarise(total_rentals = n(), 
            avg_trip_duration = mean(trip_duration),
            max_duration = max(trip_duration),
            min_duration = min(trip_duration))

user_type_summary
```


```{r}
# trip duration based on different bike types

bike_type_summary <- trip_data %>%
  group_by(rideable_type) %>%
  summarise(total_rentals = n(), 
            avg_trip_duration = mean(trip_duration),
            max_duration = max(trip_duration),
            min_duration = min(trip_duration))

bike_type_summary
```


```{r}
trip_data %>% filter(trip_duration <= 1)
```


```{r}
trip_data <- trip_data %>%
  filter(trip_duration > 1)
```


```{r}

# Filter unique start_stations
start_stations <- trip_data %>%
  distinct(start_station_id, .keep_all = TRUE) %>%
  select(start_station_id,start_lat,start_lng) %>%
  rename('station_id' = 'start_station_id',
         'lat' = 'start_lat',
         'lng' = 'start_lng')
                 
# Filter unique end_stations
end_stations <- trip_data %>%
  distinct(end_station_id, .keep_all = TRUE) %>%
  select(end_station_id,end_lat,end_lng)%>%
  rename('station_id' = 'end_station_id',
         'lat' = 'end_lat',
         'lng' = 'end_lng')


#start_stations
#end_stations

stations <- bind_rows(start_stations,end_stations)


# Remove duplicate rows based on all columns
stations_data <- distinct(stations, .keep_all = TRUE)
#stations_data

# Export the station data as a .csv file
write_csv(station_data,"../data/stations_data.csv")

```



```{r}

trips_data <- trip_data

write_csv(tripsData,"../data/trips_data.csv")

```




1.What are the most common destination of the trip starting from each station?

```{r}
# Group by start_station_id and end_station_id, and summarize the number of trips 

common_destinations <- trip_data %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) #select the most common destination for each start station

# Arrange the result in decreasing order of the number of trips 

common_destinations <- common_destinations %>%
  arrange(desc(number_of_trips))

common_destinations
```



2. Does it change with the time ?

```{r}
# Group by hour, start_station_id, and end_station_id and summarize the trip

destinations_by_hour <- trip_data %>%
  group_by(hour,start_station_id,end_station_id) %>%
  summarise(number_of_trips = n(),.groups = 'drop') %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id,hour) %>%
  slice(1) # select the most common destination for each start station at each hour

destinations_by_hour
```

3. Does it change with user type ?

```{r}
# Separate tables for member and casual users

destinations_member <- trip_data %>% 
  filter(member_casual == "member") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and user type 

# Arrange the result in decreasing order of the number of trips 
destinations_member <- destinations_member %>%
  arrange(desc(number_of_trips))

destinations_member

```

```{r}
destinations_casual <- trip_data %>% 
  filter(member_casual == "casual") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) #

# Arrange the result in decreasing order of the number of trips 
destinations_casual <- destinations_casual %>%
  arrange(desc(number_of_trips))

destinations_casual
```



4. Does it change with day of the week ?

```{r}
# Group by day_of_week, start_station_id, and end_station_name, and summarize the number of trips
destinations_by_day <- trip_data %>%
  group_by(day_of_week, start_station_id, end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id,day_of_week) %>%
  slice(1)  # Select the most common destination for each start station on each day

# View the result
destinations_by_day
```


```{r}
# Separate tables for days of week 

# Group by start_station_id, and end_station_name, and summarize the number of trips
destinations_Monday <- trip_data %>% 
  filter(day_of_week == "Monday") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and day of week


# Arrange the result in decreasing order of the number of trips 
destinations_Monday <- destinations_Monday %>%
  arrange(desc(number_of_trips))

destinations_Monday
```




```{r}
destinations_Tuesday <- trip_data %>% 
  filter(day_of_week == "Tuesday") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and day of week

# Arrange the result in decreasing order of the number of trips 
destinations_Tuesday <- destinations_Tuesday %>%
  arrange(desc(number_of_trips))

destinations_Tuesday
```



```{r}
destinations_Wednesday <- trip_data %>% 
  filter(day_of_week == "Wednesday") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and day of week

# Arrange the result in decreasing order of the number of trips 
destinations_Wednesday <- destinations_Wednesday %>%
  arrange(desc(number_of_trips))

destinations_Wednesday
```



```{r}
destinations_Thursday <- trip_data %>% 
  filter(day_of_week == "Thursday") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and day of week

# Arrange the result in decreasing order of the number of trips 
destinations_Thursday <- destinations_Thursday %>%
  arrange(desc(number_of_trips))

destinations_Thursday
```



```{r}
destinations_Friday <- trip_data %>% 
  filter(day_of_week == "Friday") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and day of week

# Arrange the result in decreasing order of the number of trips 
destinations_Friday <- destinations_Friday %>%
  arrange(desc(number_of_trips))

destinations_Friday
```


```{r}
destinations_Saturday <- trip_data %>% 
  filter(day_of_week == "Saturday") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and day of week

# Arrange the result in decreasing order of the number of trips 
destinations_Saturday <- destinations_Saturday %>%
  arrange(desc(number_of_trips))

destinations_Saturday
```



```{r}
destinations_Sunday <- trip_data %>% 
  filter(day_of_week == "Sunday") %>%
  group_by(start_station_id,end_station_id) %>%
  summarise(number_of_trips = n()) %>%
  arrange(desc(number_of_trips)) %>%
  group_by(start_station_id) %>%
  slice(1) # select the most common destination for each start station and day of week

# Arrange the result in decreasing order of the number of trips 
destinations_Sunday <- destinations_Sunday %>%
  arrange(desc(number_of_trips))

destinations_Sunday
```




For station 5905.14, the most common destination is also 5905.14, with 118 trips.

No trips start from station 5905.14 at hours 3, 4, 5, 6. At hours 0, 1, 7, 10, 12, 13, 14, 15, 16, 17, 18, 19, 21, 22, 23, the most common destination is also 5905.14. At hours 2, 11, and 20, the most common destination is 5980.07. At hour 8, the most common destination is 6441.01 with 4 trips. At hour 9, the most common destination is 5500.02 with 1 trip. Therefore, for station 5905.14, the most common destination changes with the hour of the day.

The most common destination does not change for station 5905.14 with user type. Both member and casual users’ most common destination is also 5905.14.

On all days, the most common destination for station 5905.14 is also 5905.14.


```{r}

```







```{r}
tmap_mode("view")
```



```{r}
#Number of Trips by Start Station in NYC


# Load a shape file of New York City

 nyc_shape <- st_read('../data/BoroughBoundaries.geojson')
 
 # Convert the data set to an 'sf' object
 bike_trips_sf <- st_as_sf(trip_data, coords = c("start_lng", "start_lat"))
 
 # Group data by start station and count the number of trips
 start_station_counts <- bike_trips_sf %>%
   group_by(start_station_id) %>%
   summarise(num_trips = n(), .groups = 'drop')
 
 # Create a map of start stations on top of NYC shape file
 start_station_map <-tm_shape(nyc_shape) +
   tm_borders(lwd = 1, col = "black") +  # Add borders to NYC areas
   tm_shape(start_station_counts) +
   tm_bubbles(size = 0.05, col = 'num_trips', palette = "viridis",border.col = "black") +
   tm_layout(title = 'Number of Trips by Start Station in NYC') +
   tm_basemap(server = "OpenStreetMap",alpha = 0.5)
 
 start_station_map
 
 tmap_save(tm=start_station_map,
         filename ="../output/start_station_map.png")

```



```{r}
# Load a shape file of New York City
 nyc_shape <- st_read('../data/BoroughBoundaries.geojson')
 
 # Filter the shape file to only include Manhattan
 Manhattan_shape <- nyc_shape %>%
   filter(boro_name == 'Manhattan')
 
 bike_trips_sf <- st_set_crs(bike_trips_sf, 4326)
 
 
 # Filter the trip data to include only trips that start or end within Manhattan
 bike_trips_manhattan <- st_intersection(bike_trips_sf, Manhattan_shape)
 
 
 # Group data by start station and count the number of trips
 start_station_counts_manhattan <- bike_trips_manhattan %>%
   group_by(start_station_id) %>%
   summarise(num_trips = n(), .groups = 'drop')
 
 # Create a map of start stations within Manhattan
 start_station_manhattan_map <- tm_shape(Manhattan_shape) +
   tm_borders(lwd = 1, col = "black") +
   tm_shape(start_station_counts_manhattan) +
   tm_bubbles(size = 0.05, col = 'num_trips',palette = "viridis", border.col = 'black') +
   tm_layout(title = 'Number of Trips by Start Station in Manhattan') +
   tm_basemap(server = "OpenStreetMap", alpha = 0.5)
 
 start_station_manhattan_map
 
 tmap_save(tm=start_station_manhattan_map,
         filename ="../output/start_station_manhattan_map.png")
```



```{r}

 # Load a shape file of New York City
 
 nyc_shape <- st_read('../data/BoroughBoundaries.geojson')
 
 # Convert the data set to an 'sf' object
 bike_trips_sf <- st_as_sf(trip_data, coords = c("end_lng", "end_lat"))
 
 # Group data by end station and count the number of trips
 end_station_counts <- bike_trips_sf %>%
   group_by(end_station_id) %>%
   summarise(num_trips = n(), .groups = 'drop')
 
 # Create a map of end stations on top of NYC shapefile
 end_station_map <- tm_shape(nyc_shape) +
   tm_borders(lwd = 1, col = "black") +  # Add borders to NYC areas
   tm_shape(end_station_counts) +
   tm_bubbles(size = 0.05, col = 'num_trips', palette = "YlGnBu", border.col = 'black') +
   tm_layout(title = 'Number of Trips by End Station in NYC') +
   tm_basemap(server = "OpenStreetMap", alpha = 0.5)
 
 end_station_map
 
 tmap_save(tm=end_station_map,
           filename ="../output/end_station_map.png")
```



```{r}
 # Load a shape file of New York City
 nyc_shape <- st_read('../data/BoroughBoundaries.geojson')
 
 # Filter the shape file to only include Manhattan
 Manhattan_shape <- nyc_shape %>%
   filter(boro_name == 'Manhattan')
 
 bike_trips_sf <- st_set_crs(bike_trips_sf, 4326)
 
 
 # Filter the trip data to include only trips that start or end within Manhattan
 bike_trips_manhattan <- st_intersection(bike_trips_sf, Manhattan_shape)
 
 
 # Group data by end station and count the number of trips
 end_station_counts_manhattan <- bike_trips_manhattan %>%
   group_by(end_station_id) %>%
   summarise(num_trips = n(), .groups = 'drop')
 
 # Create a map of end stations within Manhattan
 end_station_manhattan_map <- tm_shape(Manhattan_shape) +
   tm_borders(lwd = 1, col = "black") +
   tm_shape(end_station_counts_manhattan) +
   tm_bubbles(size = 0.05, col = 'num_trips',palette = "YlGnBu", border.col = 'black') +
   tm_layout(title = 'Number of Trips by End Station in Manhattan') +
   tm_basemap(server = "OpenStreetMap", alpha = 0.5)
 
 end_station_manhattan_map
 
 tmap_save(tm = end_station_manhattan_map,
         filename ="../output/end_station_manhattan_map.png")
```



```{r}
# Load a shape file of New York City
 nyc_shape <- st_read('../data/BoroughBoundaries.geojson')
 
 # Filter the shape file to only include Manhattan
 Manhattan_shape <- nyc_shape %>%
   filter(boro_name == 'Manhattan')
 
 bike_trips_sf <- st_set_crs(bike_trips_sf, 4326)
 
 
 # Filter the trip data to include only trips that start or end within Manhattan
 bike_trips_manhattan <- st_intersection(bike_trips_sf, Manhattan_shape)
 
 
# Calculate the mean trip duration for each start station 
 start_station_mean_duration <- bike_trips_manhattan %>%
   group_by(start_station_id) %>%
   summarise(mean_duration = mean(trip_duration), .groups = 'drop')
 
 # Create a map of start stations within Manhattan
 start_station_mean_duration_manhattan_map <- tm_shape(Manhattan_shape) +
   tm_borders(lwd = 1, col = "black") +
   tm_shape(start_station_mean_duration) +
   tm_bubbles(size = 0.05, col = 'mean_duration',palette = "YlGnBu", border.col = 'black') +
   tm_layout(title = 'Mean Trip Duration by Start Station in Manhattan') +
   tm_basemap(server = "OpenStreetMap", alpha = 0.5)
 
 start_station_mean_duration_manhattan_map 
 
 tmap_save(tm = start_station_mean_duration_manhattan_map,
         filename ="../output/start_station_mean_duration_manhattan_map.png")
```



```{r}
# Load a shape file of New York City
 nyc_shape <- st_read('../data/BoroughBoundaries.geojson')
 
 # Filter the shape file to only include Manhattan
 Manhattan_shape <- nyc_shape %>%
   filter(boro_name == 'Manhattan')
 
 bike_trips_sf <- st_set_crs(bike_trips_sf, 4326)
 
 
 # Filter the trip data to include only trips that start or end within Manhattan
 bike_trips_manhattan <- st_intersection(bike_trips_sf, Manhattan_shape)
 
 
# Calculate the mean trip duration for each end station 
 end_station_mean_duration <- bike_trips_manhattan %>%
   group_by(end_station_id) %>%
   summarise(mean_duration = mean(trip_duration), .groups = 'drop')
 
 # Create a map of end stations within Manhattan
 end_station_mean_duration_manhattan_map <- tm_shape(Manhattan_shape) +
   tm_borders(lwd = 1, col = "black") +
   tm_shape(end_station_mean_duration) +
   tm_bubbles(size = 0.05, col = 'mean_duration',palette = "YlGnBu", border.col = 'black') +
   tm_layout(title = 'Mean Trip Duration by End Station in Manhattan') +
   tm_basemap(server = "OpenStreetMap", alpha = 0.5)
 
 end_station_mean_duration_manhattan_map 
 
 tmap_save(tm = end_station_mean_duration_manhattan_map,
         filename ="../output/end_station_mean_duration_manhattan_map.png")
```





```{r}

```
